---
title: "Meeting 3 Analysis Script"
author: "Kenzie Nguyen"
date: "`r Sys.Date()`"
output: 
  html_document: 
    toc: true
    toc_float: true # table of contents
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(reshape2)
```

```{r load}
### Load in the data ### 

data <- read.csv("Meeting 2 Processed Data.csv")

# Recode variable

data$PIN <- as.factor(data$PIN) # Discrete categories
data$gender <- as.factor(data$gender)
data$shock_cause <- as.factor(data$shock_cause)
data$shock <- as.factor(data$shock)
data$moral <- as.factor(data$moral)

```

## Descriptives

Descriptive statistics for guilt:
```{r descriptives_guilt}

c(summary(data$guilt),"SD" = sd(data$guilt, na.rm = T)) # remove NA

```

Descriptive statistics for generosity:
```{r descriptives_generosity}

# Pre-shock
c(summary(data$preShare),"SD" = sd(data$preShare, na.rm = T)) # remove NA

# Post-shock
c(summary(data$postShare),"SD" = sd(data$postShare, na.rm = T)) # remove NA
```

Descriptive statistics for personality:
```{r descriptives_personality}

# Agreeableness
c(summary(data$agree),"SD" = sd(data$agree, na.rm = T)) # remove NA

# Neuroticism
c(summary(data$neuro),"SD" = sd(data$neuro, na.rm = T)) # remove NA
```




## Guilt Analysis

### Predicition 1.1

We expected people would feel more guilt when they caused the shock vs. their partner. 

This prediction was supported. People felt guilt when they caused a shoock to their partner.
```{r P1.1}

# Run the t-test for prediction 1.1
p1.1 <- t.test(data$guilt ~ data$shock_cause)

#Output the result
p1.1
```

There was a significant difference between the two means, **t** = `r p1.1$statistic`



### Predicition 1.2

We expected that people would feel more guilt when their partner was morally good vs. morally bad. 

This prediction was supported. people felt more guilt their partner was morally good. 
```{r p1.2}

# Run the t-test for prediction 1.1
p1.2 <- t.test(data$guilt ~ data$moral)

#Output the result
p1.2
```



### Predicition 1.3

We expected that people would feel more guilt when the partner received a stronger shock.

This prediction was supported. Guilt increased with the increasing shock magnitude. There is some increase between the different categories. 
```{r p1.3}

# Run the ANOVA for prediction 1.3
p1.3 <- lm(guilt ~ shock, data=data) # lm = linear model, ANOVA

# Output the result
anova(p1.3)

# Output the group means
tapply(data$guilt, data$shock, function(x) mean(x, na.rm = T))
```



### Prediction 2

We expected that women will feel more guilt than men. 

In spite of prior expectations, our prediction was not supported. There was no significant difference in guilt between men and women. 
```{r p2}

# Run the t-test for prediction 1.1
p2 <- t.test(data$guilt ~ data$gender)

#Output the result
p2
```



### Predicition 3

We expected agreeable people would report more guilt. 

This prediction was supported. Agreeableness was weakly, yet significantly correlated with reported levels of guilt. 
```{r p3}

# Run the correlation test for prediction 3
p3 <- cor.test(data$agree, data$guilt)

#Output the result
p3
```



### Predicition 4

We expected neurotic people would report more guilt. 

This prediction was not supported. There was no relationship between neuroticism and guilt. 
```{r p4}

# Run the correlation test for prediction 4
p4 <- cor.test(data$neuro, data$guilt)

#Output the result
p4

# Create a regression analysis of both neuro and agree as predictors for guilt
guilt_reg <- lm(guilt ~ neuro + agree, data=data) 

# linear model is the same as ANOVA and regression. Regression slope shows a similar positive/negative relationship as correlation

# Output regression result
summary(guilt_reg)
```



## Generosity 

```{r melt}

# Melt the data for within-subjects analysis

data_melt <- melt(data,
                  measure.vars = c("preShare", "postShare"),
                  variable.name = c("pre_post"),
                  value.name = c("share")
     )
```



### Predicition 5.1

We expected that people would be more generous when they caused the shock vs. when the partner caused the shock to themselves. 

This prediction was supported. Participants were more generous to their partner after the shock when the participant thought they were responsible for the shock. 
```{r p5.1}

# Perform the within-subjects ANOVA for prediction 5.1
p5.1 <- aov(share ~ shock_cause*pre_post + Error(PIN), data=data_melt)

# Output the ANOVA results:
summary(p5.1)

# Output the means separated by conditions
tapply(data_melt$share, data_melt[, c("shock_cause", "pre_post")],
       function(x) mean(x, na.rm = T))
```



### Prediction 5.2

We expected that people would be more generous when their partner was morally good vs. morally bad.

This prediction was supported. Participants were more generous to their morally good partner after the shock.
```{r p5.2}

# Perform the within-subjects ANOVA for prediction 5.1
p5.2 <- aov(share ~ moral*pre_post + Error(PIN), data=data_melt)

# Output the ANOVA results:
summary(p5.2)

# Output the means separated by conditions
tapply(data_melt$share, data_melt[, c("moral", "pre_post")],
       function(x) mean(x, na.rm = T))
```



### Predicition 5.3

We expected that people would be more generous when the shock was stronger. 

This prediction was supported. Pre-shock there was no difference in generosity, but generosity tracks level of shock 
```{r p5.3}

# Perform the within-subjects ANOVA for prediction 5.1
p5.3 <- aov(share ~ shock*pre_post + Error(PIN), data=data_melt)

# Output the ANOVA results:
summary(p5.3)

# Output the means separated by conditions
tapply(data_melt$share, data_melt[, c("shock", "pre_post")],
       function(x) mean(x, na.rm = T))
```



### Predicition 6

We expected that women would be more generous than men.

This prediction was not supported. There was not a significant interaction between gender and generosity. 
```{r p6}

# Perform the within-subjects ANOVA for prediction 5.1
p6 <- aov(share ~ gender*pre_post + Error(PIN), data=data_melt)

# Output the ANOVA results:
summary(p6)

# Output the means separated by conditions
tapply(data_melt$share, data_melt[, c("gender", "pre_post")],
       function(x) mean(x, na.rm = T))
```



### Predicition 7

We expected that agreeable people would be more generous than disagreeable people. 

This prediction was partially supported. Agreeable people were more generous after the shock but not before. 
```{r p7}

# Perform correlation test for pre shock generosity
p7pre <- cor.test(data$preShare, data$agree)

# Output the results
p7pre

# Perform correlation test for post shock generosity
p7post <- cor.test(data$postShare, data$agree)

# Output the results
p7post
```



### Predicition 8

We expected that neurotic people would be more generous than emotionally stable people.  

This prediction was not supported. There is no correlation between neuroticism and generosity before or after the shock. 
```{r p8}

# Perform correlation test for pre shock generosity
p8pre <- cor.test(data$preShare, data$neuro)

# Output the results
p8pre

# Perform correlation test for post shock generosity
p8post <- cor.test(data$postShare, data$neuro)

# Output the results
p8post
```